\Large\bf FlexibleSUSY --- A spectrum generator generator for supersymmetric models

### 2014-05-10 doc/paper.tex
improve rest of Section 6 a bit

%macros
\newcommand{\sarah}{SARAH\xspace}
\newcommand{\fs}{FlexibleSUSY\xspace}
\newcommand{\ESSM}{E$_6$SSM\xspace}
\newcommand{\code}[1]{\lstinline|#1|}  % inline source code
\newcommand{\textoverline}[1]{$\overline{\mbox{#1}}$}
\newcommand{\DRbar}{\textoverline{DR}\xspace}
\newcommand{\MSbar}{\textoverline{MS}\xspace}
\newcommand{\unit}[1]{\,\text{#1}}      % units
\newcommand{\userinput}{\text{<input>}}
\newcommand{\pole}{\text{pole}}
\newcommand{\Lagr}{\mathcal{L}}
\newcommand{\unity}{\mathbf{1}}
\newcommand{\figref}[1]{\figurename~\ref{#1}}
\newcommand{\secref}[1]{Sec.~\ref{#1}}
\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\re}{Re}
\DeclareMathOperator{\im}{Im}
\def\at{\alpha_t}
\def\ab{\alpha_b}
\def\as{\alpha_s}
\def\atau{\alpha_{\tau}}
\def\oat{O(\at)}
\def\oab{O(\ab)}
\def\oatau{O(\atau)}
\def\oatab{O(\at\ab)}
\def\oatas{O(\at\as)}
\def\oabas{O(\ab\as)}
\def\oatababq{O(\at\ab + \ab^2)}
\def\oatqatababq{O(\at^2 + \at\ab + \ab^2)}
\def\oatasatq{O(\at\as + \at^2)}
\def\oatasabas{O(\at\as +\ab\as)}
\def\oatasabasatq{O(\at\as + \at^2 +\ab\as)}
\def\oatq{O(\at^2)}
\def\oabq{O(\ab^2)}
\def\oatauq{O(\atau^2)}
\def\oabatau{O(\ab \atau)}
\def\oas{O(\as)}
\def\oatauqatab{O(\atau^2 +\ab \atau )}

@@

At the low-scale constraint the expressions for the Yukawa coupling
matrices $y_f(M_Z)$ are calculated from the tree-level mass matrices
of the SM-fermions $m_u$, $m_d$, $m_e$.  In the MSSM, for example,
these relations read

@@

At the low-scale the fermion mass matrices are composed as

@@

  m_u =
  \begin{pmatrix}
    m_{u}^{\userinput} & 0 & 0 \\
    0 & m_{c}^{\userinput} & 0 \\
    0 & 0 & m_{t,\text{susy}}^{\text{\DRbar}}(M_Z)
  \end{pmatrix} ,\\
  m_d =
  \begin{pmatrix}
    m_{d}^{\userinput} & 0 & 0 \\
    0 & m_{s}^{\userinput} & 0 \\
    0 & 0 & m_{b,\text{susy}}^{\text{\DRbar}}(M_Z)
  \end{pmatrix} ,\\
  m_e =
  \begin{pmatrix}
    m_{e}^{\userinput} & 0 & \\
    0 & m_{\mu}^{\userinput} & 0 \\
    0 & 0 & m_{\tau,\text{susy}}^{\text{\DRbar}}(M_Z)
  \end{pmatrix},

@@

where $m_{u,c,d,s,e,\mu}^{\userinput}$ are read from the \code{Block
  SMINPUTS} of the SLHA input file \cite{Skands:2003cj}.  The 3rd

@@

$m_{b,\text{SM}}^{\text{\MSbar}}(M_Z)$,

@@

detail, the top quark \DRbar mass is calculated via

@@

    \re\Sigma_{tt}^{S,\text{heavy}}(m_t^\text{pole}) \\

@@

    \left[ \re\Sigma_{tt}^{L,\text{heavy}}(m_t^\text{pole}) +
      \re\Sigma_{tt}^{R,\text{heavy}}(m_t^\text{pole}) + \Delta

@@

where the $\Sigma_{ff}^\text{heavy}$ denotes the self-energy of
fermion $f$ without the Standard Model contributions.  The appearing
QCD corrections $\Delta m_t^{(1),\text{qcd}}$ and $\Delta
m_t^{(2),\text{qcd}}$ are taken from
\cite[Eq.\ (58),(61)]{Bednyakov:2002sf} and read

@@

The \DRbar mass of the bottom quark is calculated via

@@

    \re\Sigma_{bb}^{S,\text{heavy}}(m_{b,\text{SM}}^\text{\MSbar})/m_b
    - \re\Sigma_{bb}^{L,\text{heavy}}(m_{b,\text{SM}}^\text{\MSbar}) -
    \re\Sigma_{bb}^{R,\text{heavy}}(m_{b,\text{SM}}^\text{\MSbar})} ,\\

@@

      g_2^2}{128 \pi^2} + \frac{13 g_Y^2}{1152 \pi^2}\right) ,

@@

where $m_{b,\text{SM}}^{\text{\MSbar}}(M_Z)$ is user input.  The
\DRbar mass of the $\tau$ is calculated via

@@

    \re\Sigma_{\tau\tau}^{S,\text{heavy}}(m_{\tau,\text{SM}}^\text{\MSbar}) \\

@@

      \re\Sigma_{\tau\tau}^{L,\text{heavy}}(m_{\tau,\text{SM}}^\text{\MSbar})

@@

      \re\Sigma_{\tau\tau}^{R,\text{heavy}}(m_{\tau,\text{SM}}^\text{\MSbar})

@@

  \left(1 - 3 \frac{g_Y^2 - g_2^2}{128 \pi^2}\right),

@@

where $m_{\tau,\text{SM}}^{\text{\MSbar}}(M_Z)$ is user input.  To
convert the fermion masses from the \MSbar to the \DRbar scheme the
Yukawa coupling conversion from \cite[Eq.~(19)]{Skands:2003cj} was
used and it was assumed that the VEV is already given in the \DRbar
scheme.

@@

In \fs the one-loop electroweak symmetry breaking conditions are
formulated as
%
\begin{align}
  0 = \frac{\partial V^\text{tree}}{\partial v_i} - t_i,
  \label{eq:one-loop-ewsb-eq}
\end{align}
%
where $V^\text{tree}$ is the tree-level Higgs potential, $v_i$ is the
VEVs corresponding to the Higgs field $H_i$ and $t_i$ is the one-loop
tadpole diagram of $H_i$.

@@

   return GSL_SUCCESS;
}
\end{lstlisting}
%
Here \code{x} is the vector of EWSB output parameters
(\code{EWSBOutputParameters}) and \code{f} is a vector which contains
the one-loop EWSB eqs.\ \eqref{eq:one-loop-ewsb-eq}.  This
\code{tadpole_equations()} function is called inside of
\code{MSSM<Two_scale>::solve_ewsb_iteratively_with} then via
%
\begin{lstlisting}[language=C++]
// initial guess
double x_init[number_of_ewsb_equations];
ewsb_initial_guess(x_init);

@@

If higher accuracy is required additional routines with higher order
corrections can be added in a very simple way.  For example in the
MSSM by default \fs adds two loop Higgs FORTRAN routines supplied by
P.~Slavich from \cite{slavich-degrassi} to add two loop corrections of
$\oatas$, $\oabas$, $\oatq$, $\oabatau$, $\oabq$, $\oatauq$ and
$\oatab$ as follows...  Something similar is done for the NMSSM, but
in this case the NMSSM $\oatas$, $\oabas$ pieces come from
\cite{Degrassi:2009yq}.  Corrections from other sources can be added
to the generated code by...

\subsection{Tree-level spectrum}
The tree-level $\overline{DR}$ masses are calculated by diagonalizing
the mass matrices returned from \code{SARAH`MassMatrix[]}.  The
numerical singular value decomposition is performed by the Eigen
library routine \code{Eigen::JacobiSVD} for matrices with less than
four rows and columns, and the Lapack routines \code{zgesvd},
\code{dgesvd} for larger matrices.  For the other types of
diagonalization, \code{Eigen::SelfAdjointEigenSolver} from Eigen is
used regardless of the matrix size.

The mass matrix $M^2$ for real scalar fiels $\phi_i$ is diagonalized
with an orthogonal matrix $O$ via
%
\begin{align}
  \Lagr_{m,\text{real scalar}}
  &= - \frac{1}{2} \phi^T M^2 \phi
  = - \frac{1}{2} (\phi^m)^T M^2_D \phi^m, \\
  \qquad M^2 &= (M^2)^T ,
  \qquad \phi^m = O \phi ,
  \qquad M^2_D = O M^2 O^T ,
  \qquad O^T O = \unity ,
\end{align}
%
where $M^2_D$ is diagonal and $\phi^m_i$ are the mass eigenstates.  In
case of complex scalar fields $\phi_i$ we use
%
\begin{align}
  \Lagr_{m,\text{complex scalar}}
  &= - \phi^\dagger M^2 \phi
  = - (\phi^m)^\dagger M^2_D \phi^m, \\
  \qquad M^2 &= (M^2)^\dagger ,
  \qquad \phi^m = U \phi ,
  \qquad M^2_D = U M^2 U^\dagger ,
  \qquad U^\dagger U = \unity .
\end{align}
%
A (possibly complex) symmetric mass matrix $Y$ for Weyl spinors
$\psi_i$ is diagonalized as
%
\begin{align}
  \Lagr_{m,\text{fermion}}^\text{symm.}
  &= - \frac{1}{2} \psi^T Y \psi + \text{h.c.}
  = - \frac{1}{2} \chi^T Y_D \chi + \text{h.c.}, \\
  \qquad Y &= Y^T ,
  \qquad Y_D = Z^* Y Z^\dagger ,
  \qquad \chi = Z \psi ,
  \qquad Z^\dagger Z = \unity ,
\end{align}
%
where $Y_D$ is diagonal and $\chi_i$ are the mass eigenstates.  The
phases of $Z$ are chosen such that all mass eigenvalues are positive.
In case of a non-symmetric mass matrix $X$ for Weyl spinors $\psi_i$
we use
%
\begin{align}
  \Lagr_{m,\text{fermion}}^\text{svd}
  &= - (\psi^-)^T X \psi^+ + \text{h.c.}
  = - (\chi^-)^T X_D \chi^+ + \text{h.c.}, \\
  \qquad \chi^+ &= V \psi^+ ,
  \qquad \chi^- = U \psi^- ,
  \qquad X_D = U^* X V^{-1} ,
  \qquad U^\dagger U = \unity = V^\dagger V ,
\end{align}
%
where we're again chosing the phases of $U$ and $V$ such that all mass
eigenvalues are positive.

\subsection{Two-scale fixed point iteration}
The default two-scale RG solver uses above beta functions and boundary
conditions to find the full set of $\overline{DR}$ masses in the model
consistent with all constraints.  It does so by iterating between the
scales of all boundary conditions, applying the constraints and
checking for convergence.  This approach is decribed in
\cite{Barger:1993gh} for the MSSM and is widely implemented in SUSY
spectrum generators.

The in more detail the two-scale algorithm used in \fs works as
follows (see \figurename~\ref{fig:two-scale-algorithm}):
%
\\Initial guess:
\begin{enumerate}
\item Guess gauge couplings $g_{1,2,3}$ at $m_Z$
\item Apply user-defined low-scale constraint (\code{LowScaleInput})
\item Guess Yukawa couplings $y_{u,d,e}$ at $m_Z$
\item Run model to the high-scale (\code{HighScaleFirstGuess})
\item Apply high-scale constraint (\code{HighScaleInput})
\item Run model to the low-scale (\code{LowScaleFirstGuess})
\item Solve EWSB eqs.\ at the tree-level
\item Calculate \DRbar\ masses
\end{enumerate}
%
Solving the renormalization group equations:
\begin{enumerate}
\item \label{rge-step-one} Run model to the low-scale (\code{LowScale})
  \begin{enumerate}
  \item Calculate \DRbar\ masses
  \item Recalculate low-scale
  \item Calculate \DRbar\ gauge and Yukawa couplings $g_{1,2,3}$, $y_{u,d,e}$
  \item Apply user-defined low-scale constraint (\code{LowScaleInput})
  \end{enumerate}
\item Run model to the high-scale (\code{HighScale})
  \begin{enumerate}
  \item Recalculate high-scale
  \item Apply high-scale constraint (\code{HighScaleInput})
  \end{enumerate}
\item Run model to the susy-scale (\code{SUSYScale})
  \begin{enumerate}
  \item Calculate \DRbar\ masses
  \item Recalculate low-scale
  \item Apply susy-scale constraint (\code{SUSYScaleInput})
  \item Solve EWSB eqs.\ at the one-loop level
  \end{enumerate}
\item If not converged yet, goto \ref{rge-step-one}
\end{enumerate}
%
Calculating the particle spectrum:
\begin{enumerate}
\item Run model to the susy-scale
\item Calculate the pole masses
\item Run model to the parameter output scale (SLHA block MODSEL 12)
\end{enumerate}
%
\begin{figure}[tbh]
  \centering
  \begin{tikzpicture}[node distance = 2cm, auto]
    \tikzstyle{block} = [rectangle, draw, text width=16em, text centered, minimum height=3em]
    \tikzstyle{arrow} = [draw, -latex, thick]
    \node[block] (guess) {Guess $g_i(M_Z)$, $y_f(M_Z)$ and soft
      parameters at \code{LowScale}};
    \node[block,below of=guess] (MZ) {Calculate $g_i(M_Z)$, $y_f(M_Z)$ and apply
      low-scale boundary conditions (\code{LowScaleInput})};
    \path[arrow] (guess) -- node {run to \code{LowScale}} (MZ);
    \node[block,below of=MZ] (MX) {Apply high-scale boundary conditions
      (\code{HighScaleInput})};
    \path[arrow] (MZ) -- node {run to \code{HighScale}} (MX);
    \node[block,below of=MX] (MS) {Apply susy-scale boundary conditions
      (\code{SUSYScaleInput}) and solve EWSB};
    \path[arrow] (MX) -- node {run to \code{SUSYScale}} (MS);
    %\path[arrow] (MS.east) -| node {run to \code{LowScale}} (MZ.east);
    \path[-latex, thick] (MS.east) edge[bend right=90] node[right] {run to \code{LowScale}} (MZ.east);
    \node[block,below of=MS] (spec) {Calculate pole masses};
    \path[arrow,dashed] (MS) -- node[text width=16em] {on convergence\\ run to \code{SUSYScale}} (spec);
  \end{tikzpicture}
  \caption{Iterative two-scale algorithm to calculate the spectrum.}
  \label{fig:two-scale-algorithm}
\end{figure}

\begin{figure}[tb]
  \centering
  \tikzumlset{fill class=white}
  \begin{tikzpicture}
    \umlclass[x=0, y=0, type=abstract]{Beta\_function}{
      -- scale : double\\
      -- loops : unsigned\\
      -- numPars : unsigned
    }{
      + \umlvirt{get() : const Eigen::ArrayXd} \\
      + \umlvirt{set(v : const Eigen::ArrayXd\&) : void}\\
      + \umlvirt{beta() : Eigen::ArrayXd}\\
      + run\_to(scale : double, eps : double) : void\\
    }
    \umlclass[x=8, y=-12, template={T}]{MSSM}{}{}
    \umlclass[x=8, y=-8, type=abstract]{Two\_scale\_model}{}{
      + \umlvirt{calculate\_spectrum() : void}\\
      + \umlvirt{run\_to(scale : double, eps : double) : void}\\
      + \umlvirt{set\_precision(precision : double) : void}
    }
    \umlclass[x=0, y=-4]{MSSM\_susy\_parameters}{}{
      + get() : const Eigen::ArrayXd \\
      + set(v : const Eigen::ArrayXd\&) : void\\
      + beta() : Eigen::ArrayXd
    }
    \umlclass[x=0, y=-8]{MSSM\_soft\_parameters}{}{
      + get() : const Eigen::ArrayXd \\
      + set(v : const Eigen::ArrayXd\&) : void\\
      + beta() : Eigen::ArrayXd
    }
    \umlclass[x=0, y=-12]{MSSM$<$Two\_scale$>$}{}{
      + calculate\_spectrum() : void\\
      + run\_to(scale : double, eps : double) : void\\
      + set\_precision(precision : double) : void
    }
   \umlinherit{MSSM\_susy\_parameters}{Beta\_function}
   \umlinherit{MSSM\_soft\_parameters}{MSSM\_susy\_parameters}
   \umlinherit{MSSM$<$Two\_scale$>$}{MSSM\_soft\_parameters}
   \umlinherit{MSSM$<$Two\_scale$>$}{Two\_scale\_model}
   \umldep[arg1={$<<$bind$>>$}, mult1={T $\rightarrow$ Two\_scale}, pos1=0.5]{MSSM$<$Two\_scale$>$}{MSSM}
  \end{tikzpicture}
  \caption{Two-scale model class hierarchy}
  \label{fig:two-scale-model-class-hierarchy}
\end{figure}

\begin{figure}[tb]
  \centering
  \tikzumlset{fill class=white}
  \begin{tikzpicture}
    \umlclass[x=0, y=0, type=abstract]{Two\_scale\_model}{}{
      + \umlvirt{calculate\_spectrum() : void}\\
      + \umlvirt{run\_to(scale : double, eps : double) : void}\\
      + \umlvirt{set\_precision(precision : double) : void}
    }
    \umlclass[x=0, y=-4]{RGFlow$<$Two\_scale$>$}{}{
      + solve() : void
    }
    \umlclass[x=8, y=0]{Constraint$<$Two\_scale$>$}{}{
      + \umlvirt{apply() : void}\\
      + \umlvirt{get\_scale() : double}\\
      + \umlvirt{set\_model(model : Two\_scale\_model*) : void}
    }
    \umlclass[x=8, y=4, template={T}]{Constraint}{}{}
    \umlclass[x=0, y=-7]{Initial\_guesser$<$Two\_scale$>$}{}{
      + \umlvirt{guess() : void}
    }
    \umlclass[x=0, y=-11, template={T}]{Initial\_guesser}{}{}
    \umlclass[x=8, y=-7]{Convergence\_tester$<$Two\_scale$>$}{}{
      + \umlvirt{accuracy\_goal\_reached() : bool}\\
      + \umlvirt{max\_iterations() : unsigned}
    }
    \umlclass[x=8, y=-11, template={T}]{Convergence\_tester}{}{}
    \umlclass[x=8, y=-4, template={T}]{RGFlow}{}{}
    \umldep[arg1={$<<$bind$>>$}, mult1={T $\rightarrow$ Two\_scale}, pos1=0.5]{RGFlow$<$Two\_scale$>$}{RGFlow}
    \umldep[mult1=1..*]{RGFlow$<$Two\_scale$>$}{Two\_scale\_model}
    \umldep[mult1=1..*]{RGFlow$<$Two\_scale$>$}{Constraint$<$Two\_scale$>$}
    \umldep{RGFlow$<$Two\_scale$>$}{Initial\_guesser$<$Two\_scale$>$}
    \umldep{RGFlow$<$Two\_scale$>$}{Convergence\_tester$<$Two\_scale$>$}
    \umldep[mult1={$<<$bind$>>$}, arg1={T $\rightarrow$ Two\_scale}, pos1=0.5]{Initial\_guesser$<$Two\_scale$>$}{Initial\_guesser}
    \umldep[mult1={$<<$bind$>>$}, arg1={T $\rightarrow$ Two\_scale}, pos1=0.5]{Convergence\_tester$<$Two\_scale$>$}{Convergence\_tester}
    \umldep[arg1={$<<$bind$>>$}, mult1={T $\rightarrow$ Two\_scale}, pos1=0.5]{Constraint$<$Two\_scale$>$}{Constraint}
  \end{tikzpicture}
  \caption{Two-scale renormalization group solver class hierarchy}
  \label{fig:two-scale-rgflow-class-hierarchy}
\end{figure}
{\color{red} copied and pasted from latest manual -- not sure how best to maintain this, could split maual up into tiny bits, but we may also want some differences?}

@@

\code{LowDiagonalizationPrecision}: This option provides the lowest
precision but is also the fastest one.  Here the one-loop mass matrix
$m_{f,1L}^\text{low}$ is calculated exactly once as
%
\begin{align}
  \forall i,j: (m_{f,1L}^\text{low})_{ij} = (m_{f,1L}(p^2 = m_{f_i} m_{f_j}))_{ij} ,
\end{align}
%
where $m_{f_i}$ is the $i$th mass eigenvalue of the tree-level mass
matrix $m_f$.  Afterwards $m_{f,1L}^\text{low}$ is diagonalized and
the eigenvalues are interpreted as pole masses $m_{f_i}^\pole$.

>>>>

%macros
\newcommand{\sarah}{SARAH\xspace}
\newcommand{\fs}{FlexibleSUSY\xspace}
\newcommand{\ESSM}{E$_6$SSM\xspace}
\newcommand{\code}[1]{\lstinline|#1|}  % inline source code
\newcommand{\textoverline}[1]{$\overline{\mbox{#1}}$}
\newcommand{\DRbar}{\textoverline{DR}\xspace}
\newcommand{\MSbar}{\textoverline{MS}\xspace}
\newcommand{\unit}[1]{\,\text{#1}}      % units
\newcommand{\userinput}{\text{<input>}}
\newcommand{\pole}{\text{pole}}
\newcommand{\Lagr}{\mathcal{L}}
\newcommand{\unity}{\mathbf{1}}
\newcommand{\figref}[1]{\figurename~\ref{#1}}
\newcommand{\secref}[1]{Sec.~\ref{#1}}
\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\re}{Re}
\DeclareMathOperator{\im}{Im}
\def\at{\alpha_t}
\def\ab{\alpha_b}
\def\as{\alpha_s}
\def\atau{\alpha_{\tau}}
\def\oat{O(\at)}
\def\oab{O(\ab)}
\def\oatau{O(\atau)}
\def\oatab{O(\at\ab)}
\def\oatas{O(\at\as)}
\def\oabas{O(\ab\as)}
\def\oatababq{O(\at\ab + \ab^2)}
\def\oatqatababq{O(\at^2 + \at\ab + \ab^2)}
\def\oatasatq{O(\at\as + \at^2)}
\def\oatasabas{O(\at\as +\ab\as)}
\def\oatasabasatq{O(\at\as + \at^2 +\ab\as)}
\def\oatq{O(\at^2)}
\def\oabq{O(\ab^2)}
\def\oatauq{O(\atau^2)}
\def\oabatau{O(\ab \atau)}
\def\oas{O(\as)}
\def\oatauqatab{O(\atau^2 +\ab \atau )}

@@

At the low-scale the \DRbar\ Yukawa coupling matrices
$y_f^{\text{\DRbar}}(M_Z)$ ($f=u,d,e$) in the susy model are
calculated automatically if the user has set $y_f$ to the value
\code{Automatic} in the \fs model file.  This is done for example in
the MSSM model file in \secref{Sec:modfile}.  In this case \fs
expresses the Yukawa couplings in terms of the \DRbar\ fermion mass
matrices $m_u$, $m_d$, $m_e$.  In the MSSM, for example, these
relations read

@@

The fermion mass matrices are composed as

@@

  m_u = \diag(m_{u}^{\userinput}, m_{c}^{\userinput}, m_{t,\text{susy}}^{\text{\DRbar}}(M_Z)) ,\\
  m_d = \diag(m_{d}^{\userinput}, m_{s}^{\userinput}, m_{b,\text{susy}}^{\text{\DRbar}}(M_Z)) ,\\
  m_e = \diag(m_{e}^{\userinput}, m_{\mu}^{\userinput}, m_{\tau,\text{susy}}^{\text{\DRbar}}(M_Z)),

@@

where $m_{u,c,d,s,e,\mu}^{\userinput}$ are read from the
\code{SMINPUTS} block of the SLHA input file \cite{Skands:2003cj}.
The CKM mixing matrix is currently assumed to be diagonal.  The 3rd

@@

$m_{b,\text{SM}}^{\text{\MSbar}}(M_Z)$ and

@@

detail, the top quark \DRbar mass is calculated as

@@

    \re\Sigma_{t}^{S,\text{heavy}}(m_t^\text{pole}) \\

@@

    \left[ \re\Sigma_{t}^{L,\text{heavy}}(m_t^\text{pole}) +
      \re\Sigma_{t}^{R,\text{heavy}}(m_t^\text{pole}) + \Delta

@@

where the $\Sigma_{f}^\text{heavy}$ is the self-energy of fermion $f$
without the Standard Model contributions.  The labels $L,R,S$ denote
the left-, right- and non-polarized part of the self-energy.  The
appearing QCD corrections $\Delta m_t^{(1),\text{qcd}}$ and $\Delta
m_t^{(2),\text{qcd}}$ are taken from \cite{Bednyakov:2002sf} and read

@@

The \DRbar mass of the bottom quark is calculated as

@@

    \re\Sigma_{b}^{S,\text{heavy}}(m_{b,\text{SM}}^\text{\MSbar})/m_b
    - \re\Sigma_{b}^{L,\text{heavy}}(m_{b,\text{SM}}^\text{\MSbar}) -
    \re\Sigma_{b}^{R,\text{heavy}}(m_{b,\text{SM}}^\text{\MSbar})} ,\\

@@

      g_2^2}{128 \pi^2} + \frac{13 g_Y^2}{1152 \pi^2}\right) .

@@

The \DRbar mass of the $\tau$ is calculated as

@@

    \re\Sigma_{\tau}^{S,\text{heavy}}(m_{\tau,\text{SM}}^\text{\MSbar}) \\

@@

      \re\Sigma_{\tau}^{L,\text{heavy}}(m_{\tau,\text{SM}}^\text{\MSbar})

@@

      \re\Sigma_{\tau}^{R,\text{heavy}}(m_{\tau,\text{SM}}^\text{\MSbar})

@@

  \left(1 - 3 \frac{g_Y^2 - g_2^2}{128 \pi^2}\right).

@@

To convert the fermion masses from the \MSbar to the \DRbar scheme the
Yukawa coupling conversion from \cite[Eq.~(19)]{Skands:2003cj} is used
and it is assumed that the VEV is defined in the \DRbar scheme.

@@

In \fs the one-loop electroweak symmetry breaking conditions are
formulated as
%
\begin{align}
  0 = \frac{\partial V^\text{tree}}{\partial v_i} - t_i,
  \label{eq:one-loop-ewsb-eq}
\end{align}
%
where $V^\text{tree}$ is the tree-level Higgs potential, $v_i$ is the
VEVs corresponding to the Higgs fields $H_i$ and $t_i$ is the one-loop
tadpole diagram of $H_i$.

@@

   return GSL_SUCCESS;
}
\end{lstlisting}
%
Here \code{x} is the vector of EWSB output parameters (defined via
\code{EWSBOutputParameters}) and \code{f} is a vector which contains
the one-loop EWSB eqs.\ \eqref{eq:one-loop-ewsb-eq}.  This
\code{tadpole_equations()} function is called inside of
\code{MSSM<Two_scale>::solve_ewsb_iteratively_with} then as
%
\begin{lstlisting}[language=C++]
// initial guess
double x_init[number_of_ewsb_equations];
ewsb_initial_guess(x_init);

@@

If higher accuracy is required additional routines with higher order
corrections can be added by setting \code{UseHiggs2LoopMSSM = True;}
in the model file.  For example in the MSSM by default \fs adds two
two-loop Higgs FORTRAN routines supplied by P.~Slavich from
\cite{Degrassi:2001yf,Brignole:2001jy,Dedes:2002dy,Brignole:2002bz,Dedes:2003km}
to add two loop corrections of $\oatas$, $\oabas$, $\oatq$,
$\oabatau$, $\oabq$, $\oatauq$ and $\oatab$.  In the NMSSM
contributions of the order $\oatas$, $\oabas$ \cite{Degrassi:2009yq}
can be added by setting \code{UseHiggs2LoopNMSSM = True;} in the model
file.

\subsection{Tree-level spectrum}
The tree-level \DRbar masses are calculated by diagonalizing the mass
matrices returned from \code{SARAH`MassMatrix[]}.  The numerical
singular value decomposition is performed by the Eigen library routine
\code{Eigen::JacobiSVD} for matrices with less than four rows and
columns, and the Lapack routines \code{zgesvd}, \code{dgesvd} for
larger matrices.  For the other types of diagonalization,
\code{Eigen::SelfAdjointEigenSolver} from Eigen is used regardless of
the matrix size.

\fs uses the following conventions for the diagonalization: A mass
matrix $M^2$ for real scalar fiels $\phi_i$ is diagonalized with an
orthogonal matrix $O$ as
%
\begin{align}
  \Lagr_{m,\text{real scalar}}
  &= - \frac{1}{2} \phi^T M^2 \phi
  = - \frac{1}{2} (\phi^m)^T M^2_D \phi^m, \\
  \qquad M^2 &= (M^2)^T ,
  \qquad \phi^m = O \phi ,
  \qquad M^2_D = O M^2 O^T ,
  \qquad O^T O = \unity ,
\end{align}
%
where $M^2_D$ is diagonal and $\phi^m_i$ are the mass eigenstates.  In
case of complex scalar fields $\phi_i$ we use
%
\begin{align}
  \Lagr_{m,\text{complex scalar}}
  &= - \phi^\dagger M^2 \phi
  = - (\phi^m)^\dagger M^2_D \phi^m, \\
  \qquad M^2 &= (M^2)^\dagger ,
  \qquad \phi^m = U \phi ,
  \qquad M^2_D = U M^2 U^\dagger ,
  \qquad U^\dagger U = \unity .
\end{align}
%
A (possibly complex) symmetric mass matrix $Y$ for Weyl spinors
$\psi_i$ is diagonalized as
%
\begin{align}
  \Lagr_{m,\text{fermion}}^\text{symm.}
  &= - \frac{1}{2} \psi^T Y \psi + \text{h.c.}
  = - \frac{1}{2} \chi^T Y_D \chi + \text{h.c.}, \\
  \qquad Y &= Y^T ,
  \qquad Y_D = Z^* Y Z^\dagger ,
  \qquad \chi = Z \psi ,
  \qquad Z^\dagger Z = \unity ,
\end{align}
%
where $Y_D$ is diagonal and $\chi_i$ are the mass eigenstates.  The
phases of $Z$ are chosen such that all mass eigenvalues are positive.
In case of a non-symmetric mass matrix $X$ for Weyl spinors $\psi_i$
we use
%
\begin{align}
  \Lagr_{m,\text{fermion}}^\text{svd}
  &= - (\psi^-)^T X \psi^+ + \text{h.c.}
  = - (\chi^-)^T X_D \chi^+ + \text{h.c.}, \\
  \qquad \chi^+ &= V \psi^+ ,
  \qquad \chi^- = U \psi^- ,
  \qquad X_D = U^* X V^{-1} ,
  \qquad U^\dagger U = \unity = V^\dagger V ,
\end{align}
%
where we're again chosing the phases of $U$ and $V$ such that all mass
eigenvalues are positive.

\subsection{Two-scale fixed point iteration}

The $\beta$-functions for the model parameters together with the
boundary conditions for a boundary value problem.  The default
two-scale RG solver tries to find the full set of model parameters
consistent with all constraints at all scales.  It does so by running
iteratively between the scales of all boundary conditions, imposing
the constraints and checking for convergence after each iteration.
This approach is decribed in \cite{Barger:1993gh} for the MSSM and is
widely implemented in SUSY spectrum generators.

The in more detail the two-scale algorithm used in \fs works as
follows, see \figref{fig:two-scale-algorithm}:
\subparagraph{Initial guess:} The RG solver starts to guess all model
parameters at the low-scale.
\begin{enumerate}
\item At the $M_Z$ scale the susy model gauge couplings $g_{1,2,3}$
  are set to the known Standard Model values (ignoring threshold
  corrections).
\item Then, the user-defined low-scale boundary conditions (defined in
  \code{LowScaleInput}) are imposed.
\item The Yukawa couplings $y_{u,d,e}$ of the susy model are set to
  the known Standard Model Yukawa couplings.
\item The susy parameters are run to the user-supplied first guess of
  the high-scale (\code{HighScaleFirstGuess}).
\item The high-scale boundary condition is imposed (defined in
  \code{HighScaleInput}).
\item All model parameters are run to the first guess of the low-scale
  (\code{LowScaleFirstGuess})
\item The EWSB eqs.\ are solved at the tree-level.
\item The \DRbar\ mass spectrum is calculated.
At this point all model parameters are set to some initial values and
a first estimation of the \DRbar\ mass spectrum is known.  Now the
actual iteration starts
\subparagraph{Fixed-point iteration:}
\begin{enumerate}
\item \label{rge-step-one} All model parameters are run to the
  low-scale (\code{LowScale})
  \begin{enumerate}
  \item The \DRbar\ mass spectrum is calculated.
  \item The low-scale is recalculated.
  \item The \DRbar\ gauge and Yukawa couplings $g_{1,2,3}(M_Z)$,
    $y_{u,d,e}(M_Z)$ of the susy model are calculated using threshold
    corrections.
  \item The user-defined low-scale constraint is imposed
    (\code{LowScaleInput})
\item Run all model parameters to the high-scale (\code{HighScale})
  \begin{enumerate}
  \item Recalculate the high-scale
  \item Apply the high-scale constraint (\code{HighScaleInput})
\item Run model parameters to the susy-scale (\code{SUSYScale})
  \begin{enumerate}
  \item Calculate the \DRbar\ mass spectrum
  \item Recalculate the susy-scale
  \item Apply the susy-scale constraint (\code{SUSYScaleInput})
  \item Solve the EWSB eqs.\ at the loop level
\item If not converged yet, goto \ref{rge-step-one}
If the fixed-point iteration has converged, all \DRbar\ model
parameters are known at all scales between \code{LowScale} and
\code{HighScale}.  In this case all model parameters are run to the
susy-scale and the pole-mass spectrum is calculated.  If the user has
chosen a specific output scale for the running \DRbar\ model
parameters by setting entry $12$ in block \code{MODSEL} in the SLHA
input file, all model parameters are run to the defined output scale.
\begin{figure}[tbh]
  \begin{tikzpicture}[node distance = 2cm, auto]
    \tikzstyle{block} = [rectangle, draw, text width=16em, text centered, minimum height=3em]
    \tikzstyle{arrow} = [draw, -latex, thick]
    \node[block] (guess) {Guess $g_i(M_Z)$, $y_f(M_Z)$ and soft
      parameters at \code{LowScale}};
    \node[block,below of=guess] (MZ) {Calculate $g_i(M_Z)$, $y_f(M_Z)$ and apply
      low-scale boundary conditions (\code{LowScaleInput})};
    \path[arrow] (guess) -- node {run to \code{LowScale}} (MZ);
    \node[block,below of=MZ] (MX) {Apply high-scale boundary conditions
      (\code{HighScaleInput})};
    \path[arrow] (MZ) -- node {run to \code{HighScale}} (MX);
    \node[block,below of=MX] (MS) {Apply susy-scale boundary conditions
      (\code{SUSYScaleInput}) and solve EWSB};
    \path[arrow] (MX) -- node {run to \code{SUSYScale}} (MS);
    %\path[arrow] (MS.east) -| node {run to \code{LowScale}} (MZ.east);
    \path[-latex, thick] (MS.east) edge[bend right=90] node[right] {run to \code{LowScale}} (MZ.east);
    \node[block,below of=MS] (spec) {Calculate pole masses};
    \path[arrow,dashed] (MS) -- node[text width=16em] {on convergence\\ run to \code{SUSYScale}} (spec);
  \end{tikzpicture}
  \caption{Iterative two-scale algorithm to calculate the spectrum.}
  \label{fig:two-scale-algorithm}
\end{figure}

@@

\begin{itemize}
\item \code{LowDiagonalizationPrecision}: This option provides the
  lowest precision but is also the fastest one.  Here the one-loop
  mass matrix $m_{f,1L}^\text{low}$ is calculated exactly once as
%
  \begin{align}
    \forall i,j: (m_{f,1L}^\text{low})_{ij} = (m_{f,1L}(p^2 = m_{f_i}
    m_{f_j}))_{ij} ,
  \end{align}
%
  where $m_{f_i}$ is the $i$th mass eigenvalue of the tree-level mass
  matrix $m_f$.  Afterwards $m_{f,1L}^\text{low}$ is diagonalized and
  the eigenvalues are interpreted as pole masses $m_{f_i}^\pole$.
